Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-03-18 WSSSPE2_report/WSSSPE2_review.tex
Section 1: minor changes.

The Second Workshop on Sustainable Software for Science: Practice and
Experiences
(WSSSPE2)\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe2/}} was
held on Sunday, 16 November 2014 in conjunction with the 2014 International
Conference for High Performance Computing, Networking, Storage and Analysis
(SC14)\footnote{\url{http://sc14.supercomputing.org}}.
WSSSPE2 followed a general initial workshop,
WSSSPE1\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe1/}}~\cite{WSSSPE1-pre-report,WSSSPE1},
that was held in 2013 in conjunction with the SC13 conference, and a focused
workshop,
WSSSPE1.1\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe1-1/}},
that was held in 2014 in conjunction with the SciPy conference.

Progress in scientific research is dependent on the quality and accessibility of
software at all levels and it is critical to address challenges related to the
development, deployment, maintenance, and overall sustainability of reusable
software as well as education around software practices. These challenges can be
technological, policy based, organizational, and educational, and are of
interest to developers (the software community), users (science disciplines),
software-engineering researchers, and researchers studying the conduct of
science (science of team science, science of organizations, science of science
and innovation policy, and social science communities). The WSSSPE1 workshop
engaged the broad scientific community to identify challenges and best practices
in areas of interest for sustainable scientific software. WSSSPE2 invites the
community to propose and discuss specific mechanisms to move towards an imagined
future practice for software development and usage in science and engineering.
The workshop includes multiple mechanisms for participation, encourages team
building around solutions, and identifies risky solutions with potentially
transformative outcomes. It strongly encourages participation of early-career
students and postdoctoral researchers.

This report extends a previous report that discussed the submission, peer-review, and peer-grouping processes in detail.~\cite{WSSSPE2-pre-report}.  It is also based on a collaborative set
of notes taken through a google doc during the workshop.~\cite{WSSSPE2-google-notes} 
Overall, the report discusses the organization work done before the workshop (\S\ref{sec:preworkshop}), the keynotes (\S\ref{sec:keynotes}), a session on defining sustainability (\S\ref{sec:defining}), the five breakout sessions---exploring sustainability (\S\ref{sec:exploring}); software development experiences (\S\ref{sec:devel}); credit \& incentives (\S\ref{sec:credit}); reproducibility \& reuse \& sharing (\S\ref{sec:reproduce}); code testing \& code review (\S\ref{sec:code_testing})---and some conclusions (\S\ref{sec:conclusions}).  The report also includes an incomplete list of attendees (Appendix \ref{sec:attendees}).

>>>>

The Second Workshop on Sustainable Software for Science: Practice and
Experiences
(WSSSPE2)\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe2/}} was
held on Sunday, 16 November 2014 in conjunction with the  International
Conference for High Performance Computing, Networking, Storage and Analysis
(SC14)\footnote{\url{http://sc14.supercomputing.org}}.
WSSSPE2 followed the model of a general initial workshop,
WSSSPE1\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe1/}}~\cite{WSSSPE1-pre-report,WSSSPE1},
which co-occurred with the SC13 conference, and a focused
workshop,
WSSSPE1.1\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe1-1/}},
which was organized in 2014 jointly with the SciPy conference.

Progress in scientific research is dependent on the quality and accessibility of
software at all levels. Hence it is critical to address challenges related to the
development, deployment, maintenance, and overall sustainability of reusable
software as well as education around software practices. These challenges can be
technological, policy based, organizational, and educational, and are of
interest to developers (the software community), users (science disciplines),
software-engineering researchers, and researchers studying the conduct of
science (science of team science, science of organizations, science of science
and innovation policy, and social science communities). The WSSSPE1 workshop
engaged the broad scientific community to identify challenges and best practices
in areas of interest to creating sustainable scientific software. WSSSPE2 invited the
community to propose and discuss specific mechanisms to move towards an imagined
future practice for software development and usage in science and engineering.
The workshop included multiple mechanisms for participation, encouraged team
building around solutions, and identified risky solutions with potentially
transformative outcomes. It strongly encouraged participation of early-career
scientists, postdoctoral researchers, and graduate students.

This report extends a previous report that discussed the submission, peer-review, and peer-grouping processes in detail.~\cite{WSSSPE2-pre-report}.  It is also based on a collaborative set
of notes taken through a google doc during the workshop~\cite{WSSSPE2-google-notes}.
Overall, the report discusses the organization work done before the workshop (\S\ref{sec:preworkshop}), the keynotes (\S\ref{sec:keynotes}), a session on defining sustainability (\S\ref{sec:defining}), the five breakout sessions---exploring sustainability (\S\ref{sec:exploring}); software development experiences (\S\ref{sec:devel}); credit \& incentives (\S\ref{sec:credit}); reproducibility, reuse, \& sharing (\S\ref{sec:reproduce}); code testing \& code review (\S\ref{sec:code_testing})---and some conclusions (\S\ref{sec:conclusions}).  The report also includes an incomplete list of attendees (Appendix \ref{sec:attendees}).
